Complete draft

### 2019-10-28 doc/point.tex
new version with fD as parameter in two bins Bug in normalization of D

We are interested in the properties of the function at its minimum, i.e. at $(\hat{\theta},\hat{p})$ occurs where $df/d\theta_\alpha=0 $ and $df/dp{_\alpha}=0$.
\begin{align}
0 & = \frac{1}{2}\left(\tr(C^{-1}S_\alpha C^{-1} C)  - (y-y_0)^T  C^{-1} S_{\alpha} C^{-1}  (y-y_0)\right) \\
& = \frac{1}{2}\left(\tr(C^{-1}S_\alpha C^{-1} (C_0 + \hat{\theta}_\beta S_\beta) )  - (y-y_0)^T  C^{-1} S_{\alpha} C^{-1}  (y-y_0)\right) \\
& =\hat{ \theta}_\beta F^{(\theta \theta)}_{\alpha \beta} +  \frac{1}{2}\left(\tr(C^{-1}S_\alpha C^{-1} C_0)  - (y-y_0)^T  C^{-1} S_{\alpha} C^{-1}  (y-y_0)\right) \\
\hat{ \theta}_\beta  & =  \frac{1}{2} F^{(\theta \theta)\, -1}_{\beta \alpha} \left( (y-y_0)^T  C^{-1} S_{\alpha} C^{-1}  (y-y_0) - \tr(C^{-1}S_\alpha C^{-1} C_0) \right) \\
0 &=  (y-\hat{p}_\beta x_\beta )^T C^{-1} x_\alpha \\
& = y^TC^{-1} x_\alpha - \hat{p}_\beta F^{(pp)}_{\alpha \beta} \\
 \hat{p}_\beta  & =   F^{(pp)\,-1}_{\beta \alpha} (y^TC^{-1} x_\alpha ).
\end{align}
\newpage
\section{Model}

>>>>

We are interested in the properties of the function at its minimum, i.e. at $(\hat{\theta},\hat{p})$ occurs where $df/d\theta_\alpha=0 $ and $df/dp{_\alpha}=0$.
\begin{align}
0 & = \frac{1}{2}\left(\tr(C^{-1}S_\alpha C^{-1} C)  - (y-y_0)^T  C^{-1} S_{\alpha} C^{-1}  (y-y_0)\right) \\
& = \frac{1}{2}\left(\tr(C^{-1}S_\alpha C^{-1} (C_0 + \hat{\theta}_\beta S_\beta) )  - (y-y_0)^T  C^{-1} S_{\alpha} C^{-1}  (y-y_0)\right) \\
& =\hat{ \theta}_\beta F^{(\theta \theta)}_{\alpha \beta} +  \frac{1}{2}\left(\tr(C^{-1}S_\alpha C^{-1} C_0)  - (y-y_0)^T  C^{-1} S_{\alpha} C^{-1}  (y-y_0)\right) \\
\hat{ \theta}_\beta  & =  \frac{1}{2} F^{(\theta \theta)\, -1}_{\beta \alpha} \left( (y-y_0)^T  C^{-1} S_{\alpha} C^{-1}  (y-y_0) - \tr(C^{-1}S_\alpha C^{-1} C_0) \right) \\
0 &=  (y-\hat{p}_\beta x_\beta )^T C^{-1} x_\alpha \\
& = y^TC^{-1} x_\alpha - \hat{p}_\beta F^{(pp)}_{\alpha \beta} \\
 \hat{p}_\beta  & =   F^{(pp)\,-1}_{\beta \alpha} (y^TC^{-1} x_\alpha ).
\end{align}

\newpage
\section{More Complete Model}

The model consists of top-level parameters
\begin{itemize}
\item $\gamma$: Growth index;
\item $\Omega_{M0}$: Mass density;
\item $b$: Host-galaxy bias;
\item $\sigma^2_{v,\text{NL}}$: Contribution of velocities due to overdensities of scales $k>k_{v,\text{max}}$;
\item $\sigma^2_{\rho,\text{NL}}$: Overdensities of scales $k>k_{\rho,\text{max}}$;
\item $\mathcal{M} = M -5\log{H_0}$: The zeropoint of the SN Hubble diagram;
\item $\sigma_M$: SN~Ia intrinsic magnitude dispersion.
\end{itemize}
For shorthand these are collectively labeled with $\theta$.

The model latent parameters are
\begin{itemize}
\item $\bar{\delta} | \gamma, \Omega_{M0}$: Fourier modes that describe the linear density field through $P_{\delta \delta}(f\sigma_8(\gamma, \Omega_{M0}))$;
\item $\delta | \bar{\delta}, \sigma^2_{\rho,\text{NL}}$: The density field including non-linear density effects;
\item $\delta_g | \delta, b$: The galaxy density field;
\item $v | \delta$: The velocity field excluding non-linear velocity effects;
\item $\{u_\alpha, \mu_\alpha + 5\log{H_0} \}| \delta_g $: Comoving coordinates for the  SNe in the universe;
\item $\Delta M_\alpha | \sigma_M$: Absolute magnitude difference from the average for SN $\alpha$;
\item $\Delta v_\alpha| \sigma^2_{v,\text{NL}}$:  Non-linear peculiar velocity for SN $\alpha$;
\end{itemize}
From these parameters we have the expected values of the SN observables
\begin{align*}
m_\alpha &=  (\mu_\alpha + 5\log{H_0})+\mathcal{M} + \Delta M_\alpha\\
z_\alpha &= f(\mu_\alpha,v_\alpha+\Delta v_\alpha).
%n_\alpha(u,m)& = n_0 \sum_\alpha \delta(m_\alpha,z_\alpha).
\end{align*}
%There is another set of latent parameters for a set of ``virtual'' SNe.
%\begin{itemize}
%\item $\{u_\alpha, \mu_\alpha + 5\log{H_0}\} | \delta_g$: True comoving coordinates for the set of SN;;
%\item $\Delta M_\alpha | \sigma_M$: Absolute magnitude difference from the average for SN $\alpha$;
%\end{itemize}
For shorthand these are collectively called $\phi$.

The data is a set of supernovae.  The data for supernova $i$ are 
\begin{itemize}
\item $\hat{u}_i $: Angular coordinates RA and Dec;
\item $\hat{m}_i $: peak magnitude (proxy for distance);
\item $\hat{z}_i $: redshift;
\end{itemize}

I have a model where the probability of the data is
\begin{align}
p(\bar{\delta}|\gamma,\Omega_{M0})  p(\delta| \bar{\delta},\sigma^2_{\rho,\text{NL}})  p(\delta_g|\delta,b) p(v|\delta)\\
p(\{u_\alpha, \mu_\alpha + 5\log{H_0}\} | \delta_g) \\
p(\{\Delta M_\alpha\} | \sigma_M) p(\{m_\alpha\} | \{\mu_\alpha + 5\log{H_0}\},\{\Delta M_\alpha\},\mathcal{M})\\
    p(\{\Delta v_\alpha\} | \sigma^2_v)
     p(\{z_\alpha\}| \{\mu_\alpha\}, v, \{\Delta v_\alpha\})\\
 %    p(\{\hat{u}_\alpha\}| \{u_\alpha\}) p(\{\hat{m}_\alpha\}| \{m_\alpha\})  \\
     p( \{\hat{u}_i\}, \{\hat{m}_i\}  | \{u_\alpha\}, \{m_\alpha\})       p(\{\hat{z}_i\}| \{z_\alpha\}).
\end{align}
The first line describes the fields, the second the SNe, the third their apparent magnitudes, the fourth apparent redshifts, and the last
connects the model observables with their experimental realizations.
The $p( \{\hat{u}_i\}, \{\hat{m}_i\}  | \{u_\alpha\}, \{m_\alpha\})$ does not depend on ordering.

The term $p(\{u_\alpha, \mu_\alpha + 5\log{H_0}\} | \delta_g)$ is the probability of the appearance of objects at certain coordinates and non-appearance elsewhere.
It is convenient to rewrite this term as
\begin{align}
p(\{u_\alpha, \mu_\alpha + 5\log{H_0}\} | \delta_g) & = p(0| \delta_g) \prod_\alpha \frac{ p(u_\alpha, \mu_\alpha + 5\log{H_0} | \delta_g)}{p(0_\alpha| \delta_g)}  \\
& = e^{-{\bar{N}}}  \prod_\alpha \frac{\bar{n}_\alpha^{n_\alpha}}{n_\alpha!}
\end{align}
where $\bar{N}$ is the expected number of discovered SN, $\bar{n}$ is the expected number of SN discovered at position $\alpha$, and
$n_\alpha$ is the number of SNe found at the same $\alpha$ position.  We expect $\bar{n} \propto \delta_g$.

The survey has a selection function that is dependent on the angular position and observed magnitude $S(u,m)$.
It is thus straightforward to include the selection when calculating $p( \{\hat{u}_i\}, \{\hat{m}_i\}  | \{u_\alpha\}, \{m_\alpha\})$.

\newpage

The average magnitude distance modulus of the bin is $\mu_\beta$ and the bin has linear radial peculiar velocity $v_\beta$.
Each SN~Ia in that space is parameterized by $\Delta M_i$ and $\Delta v_i$.  We assume RA and Dec are measured
perfectly, e.g. no lensing.
\section{Model}
